EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-05-14 01_intro.tex
Updated text: * Revised Repulsive Hubbard * Revised Attractive Hubbard * Introduction draft * Conclusions draft
Converging...

\section{Introduction and Motivation}\label{SecIntro}

A few words about the motivations who led us to develop this
software, possible applications and advantages:

* flexibility: can address generic cases including multi-orbital,
superconducting or spin-non-conserving regimes

* zero and low finite temperatures

* direct access to (well approximated) analytic dynamical functions

* direct access to impurity Fock space quantities

bla bla bla

>>>>

\documentclass[edipack2.tex]{subfiles}
\begin{document}

\section{Introduction and Motivation}\label{SecIntro}
{\color{red} A tentative Introduction scheme}

\paragraph{intro} Quantum impurity models play a central role in condensed matter
physics, providing the theoretical foundation for understanding
of strongly correlated electrons systems.
These models describe a small
number of interacting electronic degrees of freedom coupled to 
a non-interacting bath.
Notably, quantum impurity problems lie at the core of dynamical mean-field theory
(DMFT). By  reducing the lattice many-body problem to a
self-consistent quantum impurity problem,  retaining essential local
correlations, the DMFT enabled a  successful effective description of the
complex physics of correlated materials, including Mott insulators,
heavy fermion compounds, and high-temperature superconductors.  
However, solving the quantum impurity problems emerging in the DMFT context accurately and efficiently remains a
critical challenge, particularly for multi-orbital systems, systems
at low temperatures or endowed with lower symmetry.

\paragraph{what is new, what is inside: modes, RDM, ineq}
  In this work we introduced EDIpack2: a flexible, high-performance numerical library designed to
address these challenges, providing an exact diagonalization solver
for generic quantum impurity problems. 
The library builds on the foundations of
its predecessor featuring massively parallel Lanczos-based algorithms while including several significant
improvements. EDIpack2 supports both zero and finite temperature
calculations for different symmetries of the problem  
within a unified framework, making it suitable for a wide
range of physical phenomena, from spin-orbit coupling effect in
quantum materials to multi-orbital superconductivity. It also provides direct
access to well-approximated dynamical correlation functions across the
entire complex plane, thus enabling calculations of spectral functions, local susceptibilities, and
other response properties.

Additionally, EDIpack2 offers access to the Fock space of
the impurity system and thus, for instance, the possibility to
evaluate the impurity reduced density matrix
(iRDM). This capability enables quantum information-inspired analyses,
providing insights into the entanglement properties, subsystem purity,
and quantum correlations of interacting electron systems. These
features are increasingly recognized as essential for understanding
the emergent behavior in correlated matter, including quantum
criticality and superconductivity.

This new version includes support for systems with inequivalent
impurities, a critical feature for studying heterostructures,
superlattices, and disordered systems, where the local environment
varies significantly across different sites. This capability makes it
possible to explore a wide range of inhomogeneous systems, including
oxide interfaces or disordered systems, where
translational symmetry is broken.

\paragraph{Interoperability}
Interoperability is a core design principle of this version of 
EDIpack. The library is implemented using modern Fortran constructs, with comprehensive
C-bindings that enable integration with other programming
languages. In particular, EDIpack2 includes native  C and C++
interface layer which enable development of third-party
extensions. A significant contribution is realized via the Python API, EDIpy2, which provides a
high-level interface for defining and solving impurity problems. The
Python bindings also enable integration with established scientific
libraries, such as NumPy and SciPy, facilitating data analysis and
rapid prototyping.

The development of such Python API unlocked construction of direct interfaces to the TRIQS
(Toolbox for Research on Interacting Quantum Systems)  as
well as to the W2Dynamics software.
This integration provides EDIpack2 users with access to the
extensive data handling and analysis tools in the TRIQS ecosystem,
while also allowing TRIQS and W2Dynamics users to benefit from the
flexibility and performance of EDIpack2. This dual functionality
extends the capabilities of these platforms, supporting
reproducibility and cross-validation of numerical results, and sets a
new standard for interoperability in quantum impurity solvers.

\paragraph{closure}
In this paper, we present the mathematical framework, implementation
details, and key features of EDIpack2, along with several illustrative
applications. We aim to demonstrate how this flexible,
high-performance solver can be used to tackle a wide range of
challenging impurity problems.

\end{document}

### 2025-05-14 05_examples.tex
Updated text: * Revised Repulsive Hubbard * Revised Attractive Hubbard * Introduction draft * Conclusions draft
Converging...

\subsection{Multi-orbital impurity with Hubbard-Kanamori interaction (TRIQS)}
\label{SecExamplesTRIQS}

The third example demonstrates how to use the edipack2triqs compatibility
layer (Sec.~\ref{sSecInteropTRIQS}) to solve a quantum system comprised by a
3-orbital correlated impurity coupled to a few non-interacting bath sites.
The interaction term of the impurity Hamiltonian has the Hubbard-Kanamori
form (\ref{Hint}).

A Python script implementing a calculation that makes use of edipack2triqs
generally begins with a few module imports.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=1, lastline=13]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}

One then proceeds to defining the system under consideration. In this case we
consider an impurity atom with three correlated orbitals and two bath states
per each impurity orbital. This information must be encoded in the fundamental
set objects that are later used to construct the solver.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=15, lastline=26]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
The next step is to define a TRIQS many-body operator expression that represents
the Hamiltonian to be diagonalized.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=28, lastline=67]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
Finally, one creates a solver object and performs the actual calculation by
calling its method {\tt solve()}. This is normally the most time- and
memory-consuming step.
\lstinputlisting[style=mypython,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=69, lastline=83]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}
The computation results are readily available as attributes of the solver
object. The following code snippet shows how to access measured expectation
values of the static observables, and how to employ the plotting framework of
TRIQS to visualize the obtained Matsubara and real-frequency impurity Green's
functions (Fig.~\ref{figEx3}).
\lstinputlisting[style=mypython,
                 mathescape=false,
                 numbers=none,
                 basicstyle={\scriptsize\ttfamily},
                 firstline=85, lastline=112]
{edipack2_examples/edipack2triqs/hubbard_kanamori.py}

\begin{figure}[t!]
    \includegraphics[width=0.5\linewidth]
        {edipack2_examples/edipack2triqs/G_iw.pdf}
    \includegraphics[width=0.5\linewidth]
        {edipack2_examples/edipack2triqs/A_w.pdf}
    \caption{\label{figEx3}%
        Imaginary part of the Matsubara Green's function (left) and the
        corresponding orbital-resolved spectral function (right) computed for a
        three-orbital impurity model with an interaction of the Hubbard-Kanamori
        type (\ref{Hint}). This illustration is produced by the edipack2triqs
        example script presented in Sec.~\ref{SecExamplesTRIQS}.
    }
\end{figure}

\subsection{Bethe lattice DMFT (w2dynamics)}\label{SecExamplesBetheDMFTw2d}
Let us revisit the example of Sec.~\ref{SecExamplesBetheDMFT} to
demonstrate usage of the w2dynamics interface how to perform such a
calculation instead. Unlike when the Fortran API is used, none of the
five steps of the iterative algorithm listed there are left to the
user to implement themselves. Instead, in addition to possibly
non-default values for \NAME input variables, the model specification
must also be added to a w2dynamics parameter file and depending on the
model, further input files might need to be provided. Further
differences are that initialization is performed in w2dynamics by
starting from a self-energy of zero or reading a self-energy value
from an old file, which results in the initial Weiss field that the
initial bath parameters are fitted to, and that automatic convergence
checking is not currently available. Finally, using the same settings
as for the Fortran example we arrive at the w2dynamics parameter file
shown in the following listing.

\begin{lstlisting}[style=fstyle,language={},numbers=none,basicstyle={\scriptsize\ttfamily}]
[General]
DOS             = Bethe           # support for the Bethe lattice is built-in
half-bandwidth  = 1               # list of half-bandwidths per orbital
NAt             = 1               # number of impurities
beta            = 1000            # inverse temperature, defines Matsubara grid spacing
mu              = 1.0             # chemical potential set to achieve half-filling
EPSN            = 0.0             # turns off filling-based chemical potential search
DMFTsteps       = 100             # given no convergence checking, we might want fewer
magnetism       = para            # symmetrize self-energies per spin
FileNamePrefix  = bethe_dmft_U2   # prefix for the output file name
fileold         = bethe_dmft*hdf5 # file to read an initial self-energy from
readold         = 0               # iteration number to read initial self-energy from, 0 turns off
mixing          = 0.5             # mixing, but mixes self-energies and not Weiss fields
mixing_strategy = linear          # linear mixing as in the Fortran example
solver          = EDIPACK         # use EDIpack2 as impurity solver, not default CTHYB

[Atoms]
[[1]]                             # one subsection per impurity
Nd              = 1               # number of orbitals
Hamiltonian     = Kanamori        # create a Hubbard-Kanamori interaction Hamiltonian
Udd             = 2.0             # equivalent to ULOC
Vdd             = 1.0             # equivalent to UST, meaningless for 1 orbital
Jdd             = 0.5             # equivalent to JH, JX and JP, also meaningless here

[QMC]                             # Parameters for some grid sizes are part of this section
Ntau            = 1024            # imaginary time grid size
Niw             = 4096            # number of positive Matsubara frequencies

[EDIPACK]                         # further non-default parameters, also taken from Fortran example
NBATH           = 7               # number of bath sites
ED_TWIN         = True            # use twin symmetry
LFIT            = 2048            # number of Matsubara frequencies used for the bath fit
LANC_NGFITER    = 500             # number of Lanczos iterations for Green's function
CG_FTOL         = 1e-10           # conjugate-gradient tolerance
CG_NITER        = 2048            # maximum number of conjugate-gradient iterations
\end{lstlisting}

With this parameter file, we can use the {\tt DMFT.py} script of
w2dynamics to perform the execution and find the results in form of
both a typical w2dynamics HDF5 output file whose name starts with {\tt
  bethe\textunderscore{}dmft\textunderscore{}U2} and a number of folders whose names start with the
name of the HDF5 output file that contain unavoidable or explicitly
requested \NAME input and output files for each individual DMFT
iteration. We can extract results from the HDF5 output file e.g. using
the {\tt hgrep} script of w2dynamics. The following listing shows how
to perform a calculation with 10 parallel MPI processes and print the
impurity self-energy result of the last DMFT iteration to the terminal
assuming that the parameter file is {\tt bethe\textunderscore{}dmft.in} in the working
directory and w2dynamics installed to {\tt /opt/w2dynamics}.

\begin{lstlisting}[style=mybash]
mpiexec -n 10 /opt/w2dynamics/DMFT.py bethe_dmft.in
python /opt/w2dynamics/hgrep latest siw-full -1
\end{lstlisting}

>>>>

\subsection{Multi-orbital Hubbard (TRIQS)}

\subsection{Some model (W2Dynamics)}

### 2025-05-14 edipack2.tex
Updated text in 03_edipack2 till 1.7 Lanczos Diagonalization.
Updates from Overleaf
Finish polishing 03_edipack text Quick review of 02_install.
Updated text: * Revised Repulsive Hubbard * Revised Attractive Hubbard * Introduction draft * Conclusions draft
Converging...

\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{makecell}
\usepackage{array}
\usepackage{stackengine}
\newlength\llength
\llength=1.38ex\relax
\usepackage[dvipsnames]{xcolor}
\usepackage{makecell}
\usepackage{subfiles}
\usepackage{comment}
\newcommand{\fixme}[1]{{\color{BrickRed} #1}}
\usepackage{xspace}

@@

\definecolor{mymauve}{rgb}{0.58, 0, 0.82}
\lstdefinestyle{mybash}{
  frame=lines,
  language=bash,
  basicstyle=\footnotesize,
  captionpos=b,
  stringstyle=\color{mymauve},%\ttfamily,
  commentstyle=\itshape,
  fontadjust=true,
  keywordstyle=\bfseries,
  % It would be nice to have 'conda' here,
  % but then 'conda-forge' gets highlighted too.
  morekeywords={*,git,mkdir,cmake,make,pip},
  mathescape,
  numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=3pt
}

@@

\usepackage{bbold}
\def\11{\mathbb{1}}
\def\00{\mathbf{0}}
\def\NAME{{\rm EDIpack2}\xspace}

@@

(for the arXiv version only... ?)

Lei ha clacsonato?

Tu ha clacsonato?

Tarapia tapioco, prematurata l'interfaccia, o scherziamo? Scusi noi siamo in Monicelli, come fosse un linguaggio esoterico basato su C++ e utilizzante la toolchain di LLVM anche per Linux e Mac Os soltanto in due, oppure in quattro anche scribai con il file sorgente {\tt hm\_bethe.mc}?

Come {\tt https://github.com/lcrippa/prematurata\_la\_dmft} per esempio. No dico, mi porga il file {\tt bagaglio.cpp}. Le vede le funzioni? Lo vede che interfacciano gli array, non supportati da Monicelli, e prematurano anche?
Ora io le direi, anche con il rispetto per l'autorità, anche solo le due parole come install Monicelli from {\tt https://github.com/esseks/monicelli.git} and run Make which compiles and links the main file {\tt hm\_bethe.mc}, per esempio. Basta così, mi seguano nel programma di test. No no, Attenzione: loop dmft completo supportato secondo la ref. \cite{EDIpack} abbia pazienza. Se no, plotta i dati, anche un pochino Green's function e Self-energy in prefettura. Senza contare che {\tt prematurata\_la\_dmft} ha perso i contatti con il tarapia tapioco, dopo.

Si farà finta di passar da bischeri.

>>>>

\usepackage[english]{babel}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{makecell}
\usepackage{array}
\usepackage{stackengine}
\newlength\llength
\llength=1.38ex\relax
\usepackage[dvipsnames]{xcolor}
\usepackage{makecell}
\usepackage{subfiles}
\usepackage{comment}
\newcommand{\fixme}[1]{{\color{BrickRed} #1}}

@@

\lstdefinestyle{mybash}{
  frame=lines,
  language=bash,
  basicstyle=\footnotesize,
  captionpos=b,
  stringstyle=\color{mymauve},%\ttfamily,
  commentstyle=\itshape,
  fontadjust=true,
  keywordstyle=\bfseries,
  morekeywords={*,git,mkdir,cmake,make},
  mathescape,
  numbers=left, numberstyle=\tiny, stepnumber=1, numbersep=3pt
}

@@

\usepackage{bbold}
\def\11{\mathbb{1}}
\def\00{\mathbf{0}}
\def\NAME{{\rm EDIpack2 }}

@@
